Effective collector efficiency of sparse and dense arrays of emergent cylindrical collectors in laminar-turbulent transitional flows

### 2020-07-16 drafts/outline.tex
working on methods

Capture efficiency ($\eta=\frac{b_c}{d_c}$) is defined as the ratio between the upstream width ($b_c$) of the streamlines that encounter a collector and the collector's own diameter (\(d_c\)), which may be represented by an effective diameter in the case of non-cylindrical collectors (Figure \ref{fig:capeff}). Effective particle capture efficiency (\(\eta\prime=p_r\frac{b_c}{d_c}\)) takes into account the fact that not all particles coming into contact with a collector will adhere to it, thus including probability of retention (\(p_r\)) as a multiplicand in its calculation. This parameter, as opposed to theoretical capture efficiency $\eta$, corresponds to the empirical estimates made by our study, and presumably any others that leave $p_r$ as an implicit factor.

\begin{figure}[htbp]
\includegraphics[width=10cm]{pics/collectorefficiency.png}
\centering
\caption{Particle capture efficiency diagram (TODO: caption)} \label{fig:capeff}
\end{figure}

Particle capture in transitionally turbulent flows (\(1<\Rey<1000\)), which occur commonly in the environments and on the length scales of aquatic macrovegetation, does not follow the analytical expressions involving Reynolds number, collector diameter, and particle diameter (\(d_p\)), derived for creeping flow. Hence, empirical estimates must be used. \citet{Palmer_2004} introduced a power law expression for estimating capture efficiency of cylindrical collectors, which takes the form \[\eta=C{\Rey_c}^{a}R^{b}\,,\] where collector Reynolds number \(\Rey_c=\frac{ud_c}{\nu}\), \(u\) is flow velocity, \(\nu\) is kinematic velocity, and \(R=\frac{d_c}{d_p}\) is the ratio between collector diameter and particle diameter (\(d_p\)). \(C\), \(a\), and \(b\) are empirically determined regression coefficients, the sign and magnitude of which vary depending on number of stems, stem surface properties, and other experimental parameters \citep[e.g.,][]{Palmer_2004, Fauria_2015}.

@@

\[\frac{d\bar{\phi}}{dt} = -[\frac{Cv_s}{h}(1-E_r) + \eta^{\prime}ud_cI_c]\bar{\phi} = -k\bar{\phi}\]

@@

The assumption that probability of retention (\(p_r\)) is 100\% when stems are coated with grease might be questionable based on our result that biofilm increased \(\eta\prime\) in comparison to greased dowels. The only other effect of biofilm that is readily apparent to me is perhaps increasing effective stem diameter (\(d_c\)). However, with our model for Reynolds number (\(\Rey_c \propto d_c\)), and both us and \citet{Fauria_2015} finding \(\eta\prime \sim -\Rey_c\), increasing stem diameter would be expected to have the opposite effect, unless I'm missing something.

>>>>

Capture efficiency ($\eta=\frac{b}{d_c}$) is defined as the ratio between the upstream width ($b$) of the streamlines that encounter a collector and the collector's own diameter ($d_c$), which may be represented by an effective diameter in the case of non-cylindrical collectors (Figure \ref{fig:capeff}). Effective particle capture efficiency ($\eta\prime=p_r\frac{b_c}{d_c}$) takes into account the fact that not all particles coming into contact with a collector will adhere to it, thus including probability of retention ($p_r$) as a multiplicand in its calculation. This parameter, as opposed to theoretical capture efficiency $\eta$, corresponds to the empirical estimates made by our study, and presumably any others that leave $p_r$ as an implicit factor.

\begin{figure}[htbp]
\includegraphics[width=10cm]{pics/collectorefficiency.png}
\centering
\caption{A diagram illustrating particle capture efficiency, based on a downward-looking perspective of a cylindrical collector. $b$ is the horizontal width of upstream flow and $d_c$ is collector diameter. Adapted from \citet{Palmer_2004}} 
\label{fig:capeff}
\end{figure}

Particle capture in transitionally turbulent flows ($1<\Rey<1000$), which occur commonly in the environments and on the length scales of aquatic macrovegetation, does not follow the analytical expressions involving Reynolds number, collector diameter, and particle diameter ($d_p$), derived for creeping flow. Hence, empirical estimates must be used. \citet{Palmer_2004} introduced a power law expression for estimating capture efficiency of cylindrical collectors, which takes the form \[\eta=C{\Rey_c}^{a}R^{b}\,,\] where collector Reynolds number $\Rey_c=\frac{ud_c}{\nu}$, $u$ is flow velocity, $\nu$ is kinematic velocity, and $R=\frac{d_c}{d_p}$ is the ratio between collector diameter and particle diameter ($d_p$). $C$, $a$, and $b$ are empirically determined regression coefficients, the sign and magnitude of which vary depending on number of stems, stem surface properties, and other experimental parameters \citep[e.g.,][]{Palmer_2004, Fauria_2015}.

@@

\[\frac{d\bar{\phi}}{dt} = -[\frac{Cv_s}{h}(1-E_r) + \eta^{\prime}ud_cI_c]\bar{\phi} = -k\bar{\phi}\] % \[ is necessary instead of $$ because it preserves line numbers

@@

The assumption that probability of retention ($p_r$) is 100\% when stems are coated with grease might be questionable based on our result that biofilm increased $\eta\prime$ in comparison to greased dowels. The only other effect of biofilm that is readily apparent to me is perhaps increasing effective stem diameter ($d_c$). However, with our model for Reynolds number ($\Rey_c \propto d_c$), and both us and \citet{Fauria_2015} finding $\eta\prime \sim -\Rey_c$, increasing stem diameter would be expected to have the opposite effect, unless I'm missing something.
